Coronal Fourier power spectra: implications for coronal seismology and coronal heating.

### 2014-06-13 tex/paper1/paper1.tex
plots, and paper edits

\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{jack.ireland@nasa.gov}
\newcommand{\PS}{power spectrum}
\newcommand{\PL}{power-law}
\newcommand{\Fps}{Fourier \PS}

@@

\begin{abstract}
The Fourier power spectra of regions of the solar corona are
investigated using AIA 171\AA\ and 193\AA\ data.  It is shown that the
Fourier power spectra in all the regions show a \PL\ behaviour
approximately.  This suggests that in many parts of the corona a
background model consisting of white noise and a trend cannot be
assumed when analyzing for the presence of oscillations in the corona.
The implications of this discovery to the field of coronal seismology
and automated detections of oscillations is discussed.  Further, it is
shown that the observed \PL\ \PS\ can be formed by summing a
distribution of exponentially decaying emission events along the line
of sight.  This is consistent with the notion that the solar
atmosphere is heated everywhere by small energy deposition events.

\end{abstract}

@@

\begin{figure}
\centerline{
\plottwo{shutdownfun3_6hr_disk_1.5_171_loopfootpoints.hanning.relative.power_spectra_distributions.eps}{shutdownfun3_6hr_disk_1.5_193_loopfootpoints.hanning.relative.power_spectra_distributions.eps}
}
\centerline{
\plottwo{shutdownfun3_6hr_disk_1.5_171_moss.hanning.relative.power_spectra_distributions.eps}{shutdownfun3_6hr_disk_1.5_193_moss.hanning.relative.power_spectra_distributions.eps}
}
\centerline{
\plottwo{shutdownfun3_6hr_disk_1.5_171_qs.hanning.relative.power_spectra_distributions.eps}{shutdownfun3_6hr_disk_1.5_193_qs.hanning.relative.power_spectra_distributions.eps}
}
\centerline{
\plottwo{shutdownfun3_6hr_disk_1.5_171_sunspot.hanning.relative.power_spectra_distributions.eps}{shutdownfun3_6hr_disk_1.5_193_sunspot.hanning.relative.power_spectra_distributions.eps}
}
\caption{Distribution of Fourier power at some sample frequencies as a
function of AIA waveband and location.}
\label{fig:dist171193}
\end{figure}
Figure \ref{fig:dist171193} show the distributions of the logarithm of
the Fourier power for a small number of frequencies, for all the
pixels in each region.  These distributions are centrally peaked, and
show a slightly heavier tail towards lower Fourier powers (negative
skew), and are slightly broader when compared to a Gaussian
distribution.  Plots of all these distribution show that they are all
unimodal.  These observations guide the choice of how to average the
\PS\ over each of the selected region.  The nature of these
distributions suggest that the arithmetic mean of the Fourier power is
biased by the very largest powers.  Therefore, the mean value of the
logarithm of the Fourier power is a better description of the range of
values in the data (this is equivalent to the geometric mean of the
Fourier power).  The \FPs\ used below are calculated from the mean
value of the logarithm of the Fourier power.

\begin{figure}
\centerline{
\plottwo{shutdownfun3_6hr_disk_1.5_171_loopfootpoints.hanning.relative.logiobs.1.model_fit_compare.pymc.eps}{shutdownfun3_6hr_disk_1.5_193_loopfootpoints.hanning.relative.logiobs.1.model_fit_compare.pymc.eps}
}
\centerline{
\plottwo{shutdownfun3_6hr_disk_1.5_171_moss.hanning.relative.logiobs.1.model_fit_compare.pymc.eps}{shutdownfun3_6hr_disk_1.5_193_moss.hanning.relative.logiobs.1.model_fit_compare.pymc.eps}
}
\centerline{
\plottwo{shutdownfun3_6hr_disk_1.5_171_qs.hanning.relative.logiobs.1.model_fit_compare.pymc.eps}{shutdownfun3_6hr_disk_1.5_193_qs.hanning.relative.logiobs.1.model_fit_compare.pymc.eps}
}
\centerline{
\plottwo{shutdownfun3_6hr_disk_1.5_171_sunspot.hanning.relative.logiobs.1.model_fit_compare.pymc.eps}{shutdownfun3_6hr_disk_1.5_193_sunspot.hanning.relative.logiobs.1.model_fit_compare.pymc.eps}
}
\caption{Geometric mean of the Fourier power spectra in each of the
  four regions studied in AIA 171\AA\ (left column) and AIA
  193\AA\ (right column).  Also shown are the posterior mean and
  maximum likelihood fits for model $M_{1}$ and $M_{2}$ (see Section
  \ref{sec:anal} and Appendix \ref{sec:app:ind} for more detail).
  Model fit parameter values and their 68\% credible intervals are
  quoted in Table \ref{tab:parameters}.}
\label{fig:fit171193}
\end{figure}

@@

The (geometric) mean Fourier power spectra are shown in Figure
\ref{fig:fit171193} for each of the physical regions considered.  It
is clear that, in general, the Fourier power spectra exhibit power-law
like characteristics.  The results for the quiet Sun, loop footpoints
and sunspot regions suggest a \PL\ \PS.  This can be modeled as
\begin{equation}
\label{eqn:pwrlaw}
\mbox{Model $M_{1}$}: P_{1}(\nu) = A\nu^{-n} + C,
\end{equation}
where $\nu$ is the normalized frequency, $A>0$ is a proportionality
constant and $n>0$ is the \PL\ index.  The quantity $C>0$ models
the high-frequency / low power end of the spectrum where the detection
properties of the detector apparatus are assumed to dominate the
observations.

The moss results for 171\AA\ and 193\AA\ are notably different from
the other results.  In comparison to the general trend observed in
other regions, there appears to be excess Fourier power in the range
1-10 mHz.  \cite{2003ApJ...595L..63D} show that TRACE
171\AA\ (TRansition Region And Coronal Explorer) time-series of bright
upper transition region emission above active region plage (also known
as moss) is a source of wavelet power. Periods of significant wavelet
power are found in the range 200 to 600 seconds, and typically persist
for 4–7 cycles.  Later work by \cite{2005ApJ...624L..61D} suggested
that spicule flux tubes tilted with respect to the solar surface
provide a mechanism by which oscillatory power from lower in the
atmosphere may be channeled to upper portions of the atmosphere.  This
suggests that a second model $M_{2}$ for the \Fps\ in regions should
be considered.  The second model $M_{2}$ has two contributions to the
overall \Fps; a background power-law of the form of $P_{1}$, and a
contribution that is more localized, $G(\nu)$.  The model is given by
\begin{equation}
\label{eqn:pwrlawbump}
\mbox{Model $M_{2}$}: P_{2}(\nu) = P_{1}(\nu) + G(\nu)
\end{equation}
where the localized contribution is described by
\begin{equation}
\label{eqn:bump}
G(\nu) = \alpha\exp\left[-\frac{(\ln(\nu)-\beta)^{2}}{2\sigma^{2}}\right].
\end{equation}
Both these models are fit to the geometrically averaged Fourier power
spectrum in both wavelengths and in all four regions, and compared to
decide which model best describes the data.  The model comparison
process is described in Section \ref{ssec:compare}.

\subsection{Comparing spectral models}\label{ssec:compare}
Deciding which model best fits the data is achieved using a model
comparison process based on the work of \cite{2010MNRAS.402..307V}.
In the first part of this work, two models of the background Fourier
power spectrum from XMM-Newton observations of highly variable Seyfert
1 galaxies are compared to find which one best describes the
background spectrum.  The comparison process is as follows.

@@

@@

Time series of AIA 171\AA\ and 193\AA images show approximately power
law like properties over all regions of the solar surface.  Detecting
narrow band oscillatory power should take in to account the presence
of the \PL\ like \PS.  The power spectra have been modelled using
three components: a \PL\ \PS, a lognormal component, and a constant
background.  It is hypothesized that the power-law component is due to
the sum along the line of sight of a large number of statistically
similar energy deposition events.  It is also hypothesized that the
lognormal component is due to the channeling of oscillatory power from
lower to higher levels of the solar atmosphere.

@@

\section{Fitting models to the \PS}\label{sec:app:ind}
In order to perform the fit, the noise in the mean \FPs\ must be
estimated.  To do that, the power distributions of Section \ref{???}
are considered.  Each measurement of the Fourier power at a given
pixel is not independent of the Fourier power at another pixel.  This
is because neighbouring pixels are highly likely to be physically
connected to each other, and so time-series in neighboring pixels are
likely to be strongly correlated (assuming that the background
structure exists over at least two pixels).  Also, the point spread
function of AIA spreads emission in one pixel over on to its
neighbors.  Secondly, the distributions are only approximately normal.

@@

Hence, the number of effectively independent pixels time-series must
be calculated.  First, a pair of randomly chosen next-nearest neighbor
pixels are selected from the region.  Next nearest neighbor pixels are
chosen as close physical proximity is most likely to express the
strongest interdependence.  Next, the cross correlation $c(\tau)$
function between two randomly chosen next-nearest neighbor pixels is
calculated.  The cross correlation coefficient measures the linear
dependence of two time series $X(t)$ and $Y(t \pm \tau)$ at different
lags $\tau$.  This is used to define a coefficient of independence
between the neighbors,
\begin{equation}
\label{eqn:ind}
\rho = 1 - \max|c(\tau)|.
\end{equation}
When $\rho$ is 1, the cross-correlation coefficient is zero, and there
is no measurable linear dependence between neighboring time series.
When $\rho$ is zero, the cross correlation coefficient is $\pm 1$, and
the two time-series are linearly dependent on each other.  This
procedure is repeated 10,000 times and the mean value $\hat{\rho}$ is
calculated. The effective number of independent pixels in region is
given by
\begin{equation}
\label{eqn:nind}
N_{eff}= 1 + \hat{\rho}(N_{pixel}-1).
\end{equation}
The number of independent measurements $N$ is set to $N_{eff}$ in
Equation \ref{eqn:sigmamean}.  The value $\sigma_{mean}$ is used as
the estimate of the error in the \FPs.  Table \ref{tab:neff} shows the
number of pixels in the region and the estimated number of effective
pixels.

@@

The \FPs\ are assumed to be normally distributed with a width given by
$\sigma_{mean}$ as calculated above.
Equations \ref{eqn:pwrlaw} and \ref{eqn:pwrlawbump} are fit to the
geometric mean power spectra shown in Figures \ref{fig:fit171193}.
The frequency $\nu$ is defined as $f/f_{0}$ where $f$ is the FFT
frequency in Hz, and $f_{0}$ is the lowest non-zero FFT frequency
defined by the data (Section \ref{sec:obs}).  A five step fitting
process is used to fit Models 1 and 2 to the data.  Maximum likelihood
estimates are found using scipy's maximum likelihood estimators

>>>>

\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{jack.ireland@nasa.gov}
\newcommand{\PS}{power spectrum}
\newcommand{\mPS}{mean power spectrum}

\newcommand{\PA}{power spectra}
\newcommand{\mPA}{mean power spectra}

\newcommand{\PL}{power-law}

\newcommand{\Fps}{Fourier \PS}
\newcommand{\mFps}{mean Fourier \PS}

\newcommand{\Fpa}{Fourier \PA}
\newcommand{\mFpa}{mean Fourier \PA}

@@

\begin{abstract}
The \mFpa\ of regions of the solar corona are investigated using AIA
171\AA\ and 193\AA\ data.  It is shown that the \mFpa\ in all regions
show an approximate \PL\ behaviour.  This average behavior is
inconsistent with the common assumption that coronal time-series are
well described by the sum of a long time-scale background trend plus
Gaussian-distributed noise, with some specific locations also showing
an oscillatory signal.  The implications of this discovery to the
field of coronal seismology and automated detections of oscillations
is discussed.  Further, it is shown that the observed \PL\ \PS\ can be
formed by summing a distribution of exponentially decaying emission
events along the line of sight.  This is consistent with the notion
that the solar atmosphere is heated everywhere by small energy
deposition events.
\end{abstract}

@@

\begin{figure}
\centerline{
\epsscale{0.8}\plottwo{shutdownfun3_6hr_disk_1.5_171_loopfootpoints.hanning.relative.power_spectra_distributions.eps}{shutdownfun3_6hr_disk_1.5_193_loopfootpoints.hanning.relative.power_spectra_distributions.eps}
}
\centerline{
\epsscale{0.8}\plottwo{shutdownfun3_6hr_disk_1.5_171_moss.hanning.relative.power_spectra_distributions.eps}{shutdownfun3_6hr_disk_1.5_193_moss.hanning.relative.power_spectra_distributions.eps}
}
\centerline{
\epsscale{0.8}\plottwo{shutdownfun3_6hr_disk_1.5_171_qs.hanning.relative.power_spectra_distributions.eps}{shutdownfun3_6hr_disk_1.5_193_qs.hanning.relative.power_spectra_distributions.eps}
}
\centerline{
\epsscale{0.8}\plottwo{shutdownfun3_6hr_disk_1.5_171_sunspot.hanning.relative.power_spectra_distributions.eps}{shutdownfun3_6hr_disk_1.5_193_sunspot.hanning.relative.power_spectra_distributions.eps}
}
\caption{Distribution of Fourier power at some sample frequencies as a
function of AIA waveband and location.}
\label{fig:dist171193}
\end{figure}
Figure \ref{fig:dist171193} show the distributions of the logarithm of
the Fourier power for a small number of frequencies, for all the
pixels in each region.  These distributions are centrally peaked, and
show a slightly heavier tail towards lower Fourier powers (negative
skew), and are slightly broader when compared to a Gaussian
distribution.  Plots of all these distribution show that they are all
unimodal.  These observations guide the choice of how to average the
\PS\ over each of the selected region.  The nature of these
distributions suggest that the arithmetic mean of the Fourier power is
biased by the very largest powers.  Therefore, the mean value of the
logarithm of the Fourier power is a better description of the range of
values in the data (this is equivalent to the geometric mean of the
Fourier power).  The \PA\ used below are calculated from the mean
value of the logarithm of the Fourier power.

\begin{figure}
\centerline{
\epsscale{0.8}\plottwo{shutdownfun3_6hr_disk_1.5_171_loopfootpoints.hanning.relative.logiobs.1.model_fit_compare.pymc.eps}{shutdownfun3_6hr_disk_1.5_193_loopfootpoints.hanning.relative.logiobs.1.model_fit_compare.pymc.eps}
}
\centerline{
\epsscale{0.8}\plottwo{shutdownfun3_6hr_disk_1.5_171_moss.hanning.relative.logiobs.1.model_fit_compare.pymc.eps}{shutdownfun3_6hr_disk_1.5_193_moss.hanning.relative.logiobs.1.model_fit_compare.pymc.eps}
}
\centerline{
\epsscale{0.8}\plottwo{shutdownfun3_6hr_disk_1.5_171_qs.hanning.relative.logiobs.1.model_fit_compare.pymc.eps}{shutdownfun3_6hr_disk_1.5_193_qs.hanning.relative.logiobs.1.model_fit_compare.pymc.eps}
}
\centerline{
\epsscale{0.8}\plottwo{shutdownfun3_6hr_disk_1.5_171_sunspot.hanning.relative.logiobs.1.model_fit_compare.pymc.eps}{shutdownfun3_6hr_disk_1.5_193_sunspot.hanning.relative.logiobs.1.model_fit_compare.pymc.eps}
}
\caption{Geometric mean of the \protect\Fpa\ in each of the
  four regions studied in AIA 171\AA\ (left column) and AIA
  193\AA\ (right column).  Also shown are the posterior mean and
  maximum likelihood fits for model $M_{1}$ and $M_{2}$ (see Section
  \ref{sec:anal} and Appendix \ref{sec:app:ind} for more detail).
  Model fit parameter values and their 68\% credible intervals are
  quoted in Table \ref{tab:parameters}.}
\label{fig:fit171193}
\end{figure}

@@

The (geometric) \mFpa\ are shown in Figure \ref{fig:fit171193} for
each of the physical regions considered.  It is clear that, in
general, the \mFpa\ exhibit power-law like characteristics.  The
results for the quiet Sun, loop footpoints and sunspot regions suggest
a \PL\ \PS.  This can be modeled as
\begin{equation}
\label{eqn:pwrlaw}
\mbox{Model $M_{1}$}: P_{1}(\nu) = A\nu^{-n} + C,
\end{equation}
where $\nu$ is the normalized frequency, $A>0$ is a proportionality
constant and $n>0$ is the \PL\ index.  The quantity $C>0$ models
the high-frequency / low power end of the spectrum where the detection
properties of the detector apparatus are assumed to dominate the
observations.

The moss results for 171\AA\ and 193\AA\ are notably different from
the other results.  In comparison to the general trend observed in
other regions, there appears to be excess Fourier power in the range
1-10 mHz.  \cite{2003ApJ...595L..63D} show that TRACE
171\AA\ (TRansition Region And Coronal Explorer) time-series of bright
upper transition region emission above active region plage (also known
as moss) is a source of wavelet power. Periods of significant wavelet
power are found in the range 200 to 600 seconds, and typically persist
for 4–7 cycles.  Later work by \cite{2005ApJ...624L..61D} suggested
that spicule flux tubes tilted with respect to the solar surface
provide a mechanism by which oscillatory power from lower in the
atmosphere may be channeled to upper portions of the atmosphere.  This
suggests that a second model $M_{2}$ for the \Fps\ in regions should
be considered.  The second model $M_{2}$ has two contributions to the
overall \Fps; a background power-law of the form of $P_{1}$, and a
contribution that is more localized, $G(\nu)$.  The model is given by
\begin{equation}
\label{eqn:pwrlawbump}
\mbox{Model $M_{2}$}: P_{2}(\nu) = P_{1}(\nu) + G(\nu)
\end{equation}
where the localized contribution is described by
\begin{equation}
\label{eqn:bump}
G(\nu) = \alpha\exp\left[-\frac{(\ln(\nu)-\beta)^{2}}{2\sigma^{2}}\right].
\end{equation}
Both these models are fit to the geometrically averaged Fourier power
spectrum in both wavelengths and in all four regions (Figure
\ref{fig:fit171193}), and compared to decide which model best
describes the data. The details of the model fitting process are
described in \ref{sec:app:ind}.

@@

Model selection is implemented using the Akaike Information Criterion
(AIC) and the Bayesian Information Criterion (BIC)
\citep{2007MNRAS.377L..74L}.  The AIC is defined as
\begin{equation}\label{eqn:aic}
AIC \equiv -2 \ln L_{max} - 2k
\end{equation}
where $L_{max}$ is the maximum likelihood achievable by the model, and
$k$ is the number of parameters in the model.  The BIC is defined as
\begin{equation}\label{eqn:bic}
BIC \equiv -2 \ln L_{max} - k\ln N
\end{equation}
where $N$ is the number of data points used in the fit.  Model
selection is guided by calculating $\delta AIC = AIC_{1} - AIC_{2}$
and $\delta BIC = BIC_{1} - BIC_{2}$.  Positive values of $\delta AIC,
\delta BIC$ express a preference for model 2 over model 1.

For both the AIC and the BIC, Model 2 is overwhelmingly preferred to
Model 1. Table \ref{tab:parameters} gives the parameter estimates and
68\% credible intervals of the parameter values.  The power law
indices for all regions and both wavebands lie in the range 1.8 to
2.3, although there is no consistency between one waveband and
another.  For example, the 171\AA\ and 193\AA\ loop footpoint results
both have the same power law index, but the other regions do not.
However, it should be noted that there is no {\it a priori} reason why
the power law indices for a region seen in two different wavebands
should be the same.  This is because each waveband is designed to
examine different features on the Sun in broadly different temperature
ranges, and may be imaging very different physical processes: hence,
the emission seen in one waveband need not have the same statistical
properties as another

@@

Time series of AIA 171\AA\ and 193\AA\ images show approximately power
law like properties over all regions of the solar surface.  Detecting
narrow band oscillatory power should take in to account the presence
of the \PL\ like \PS.  The power spectra have been modelled using
three components: a \PL\ \PS, a lognormal component, and a constant
background.  It is hypothesized that the power-law component is due to
the sum along the line of sight of a large number of statistically
similar energy deposition events.  It is also hypothesized that the
lognormal component is due a combination of the channeling of
oscillatory power from lower to higher levels of the solar atmosphere
and lower temperature emission lines in the AIA 171\AA\ and
193\AA\ bandpasses.

@@

\section{Fitting models to the \mFps}\label{sec:app:ind}

In order to perform the fit, the noise in the \mFps\ must be
estimated.  To do that, the power distributions of Section \ref{???}
are considered.  The Fourier power at a given pixel is not independent
of the Fourier power at another pixel.  This is because neighbouring
pixels are highly likely to be physically connected to each other, and
so time-series in neighboring pixels are likely to be strongly
correlated (assuming that the background structure exists over at
least two pixels).  Also, the point spread function of AIA spreads
emission in one pixel over on to its neighbors.  Secondly, the
distributions are only approximately normal.

@@

Hence, the number of effectively independent pixels time-series must
be calculated.  First, a pair of randomly chosen next-nearest neighbor
pixels are selected from the region.  Next nearest neighbor pixels are
chosen as close physical proximity is most likely to express the
strongest interdependence.  Next, the cross correlation $c(\tau)$
function between two randomly chosen next-nearest neighbor pixels is
calculated.  The cross correlation coefficient measures the linear
dependence of two time series $X(t)$ and $Y(t \pm \tau)$ at different
lags $\tau$.  This is used to define a coefficient of independence
between the neighbors,
\begin{equation}
\label{eqn:ind}
\rho = 1 - \max|c(\tau)|.
\end{equation}
When $\rho$ is 1, the cross-correlation coefficient is zero, and there
is no measurable linear dependence between neighboring time series.
When $\rho$ is zero, the cross correlation coefficient is $\pm 1$, and
the two time-series are linearly dependent on each other.  This
procedure is repeated 10,000 times and the mean value $\hat{\rho}$ is
calculated. The effective number of independent pixels in region is
given by
\begin{equation}
\label{eqn:nind}
N_{eff}= 1 + \hat{\rho}(N_{pixel}-1).
\end{equation}
The number of independent measurements $N$ is set to $N_{eff}$ in
Equation \ref{eqn:sigmamean}.  The value $\sigma_{mean}$ is used as
the estimate of the error in the \Fps.  Table \ref{tab:neff} shows the
number of pixels in the region and the estimated number of effective
pixels.

@@

The \mFpa\ are assumed to be normally distributed with a width given by
$\sigma_{mean}$ as calculated above.
Equations \ref{eqn:pwrlaw} and \ref{eqn:pwrlawbump} are fit to the
geometric mean power spectra shown in Figures \ref{fig:fit171193}.
The frequency $\nu$ is defined as $f/f_{0}$ where $f$ is the FFT
frequency in Hz, and $f_{0}$ is the lowest non-zero FFT frequency
defined by the data (Section \ref{sec:obs}).  A five step fitting
process is used to fit Models 1 and 2 to the data.  Maximum likelihood
estimates are found using scipy's maximum likelihood estimators
